Open science and the future university researcher\thanks{This work is licensed under a \href{https://creativecommons.org/licenses/by/4.0/}{Creative Commons Attribution 4.0 International License} \ccby}

### 2017-02-07 paper-imagine-new-university.tex
edits before final version

\begin{abstract}
\noindent
This white paper was written as a contribution to the ``Imagining Tomorrow's University:
Rethinking scholarship, education, and institutions for an open, networked era''
workshop, a joint NIH\slash NSF-funded event held 8--9 March 2017 in Rosemont, IL.
In this paper, I present an overview of what I consider open science,
why it is important, and how it plays a role in my research agenda.
I also discuss challenges faced in pursuing research openness.
This views are based on my experiences and plans. As such, all views expressed
here are personal, and do not necessarily represent those
of my institution or colleagues.
\end{abstract}

@@

Transforming research communities from traditional, closed environments to open
ones is important for a number of reasons, including (but not necessarily limited
to) the following six. McKiernan et al.\autocite{McKiernan:2016iz} discuss these
and additional benefits for researchers working openly. Tennant et al.\autocite{Tennant:2016bi}
review in detail the benefits of open-access publications.

\paragraph{Accessibility:}
Openness in research ensures that research products,
particularly written output, remain accessible to the public. This includes the
research community, funders, policy makers, and the general public. Accessibility
of research products is particularly important for publicly funded
research---since the public paid for the research, the public should have access
and be able to benefit from it.
(This does not prevent innovators or other parties from developing commercial
intellectual property based on the findings, but ensures that the original
discovery, when funded by the public, remains accessible to all.)

\paragraph{Reproducibility:}
Releasing products of research, including software and data, helps enable
reproducibility. This is particularly true for computational science, where a
written description of methods can never describe an approach as completely as
source code. In general, access to research software used to perform a
computational study, or the data from an experimental study, should enable
others to reproduce the findings of the original researchers.
However, open science is a necessary but not sufficient aspect of reproducibility,
as it can be challenging to reproduce\slash replicate results even with
available research software and data\autocite{Mesnard:2016,Barba:2016ky}.

@@

\paragraph{It's nice:}
In addition to the above benefits, sharing products of research openly is kind
to colleagues and the greater research community, as it prevents people from
wasting time repeating work unnecessarily.
Like many graduate students, I began working on my dissertation research by
attempting to reimplement another group's method and reproduce some of the
results in a paper. However, this group did not share any source code openly.
As a result, I wasted significant time dealing with minor implementation details
or inputs not discussed in their papers. This could have been avoided by sharing
the source code, which would have allowed me to more quickly move on to new work.
Graduate students and other researchers constantly face similar challenges that
could be avoided by greater openness in research.

@@

Many published journal articles go unread, even in their topical domains.
One study of citation rates found that 27\% of papers published in the natural
sciences and engineering go uncited\autocite{Lariviere:2009}\footnote{Of course,
papers that are read may not be cited, and papers that are cited may not actually
be read.}
Those who do read most papers likely come from research institutions similar to
those of the authors, even if the findings could be impactful beyond these
confines, for example leading to policy changes or technological solutions
for the developing world. In part this is due to the challenging technical content,
jargon, and niche topics---but it is also due to the lack of access to the journals
where most research findings reside. (Making the content of these papers actually
understandable or digestible by most is another challenge.)

Considering the high and ever-increasing cost of scholarly journal subscriptions,
research results should not be limited to those with the means to purchase access.
By publishing articles in open-access journals or self-archiving (i.e., green open
access), researchers can ensure access for all more members of society, including
policymakers, funders, members of the media, entrepreneurs, and the general
public---or scientists and engineers in the Global South.

@@

Over the past year and a half (September 2015--February 2017), as I initiated
and built my research group, I worked to crystallize my vision and plan for
performing research openly. The guiding principles I formed were heavily
inspired by the Reproducibility PI Manifesto of Lorena Barba\autocite{Barba:2012pi},
the Peer Reviewers' Openness Initiative\autocite{Morey150547}, and others.
However, in addition to these exemplars, my goals are also motivated by the desire
to work more openly than the norm in my field.

My field, which includes computational modeling of combustion and fluid dynamics,
chemical kinetics, and associated numerical methods, lacks reputable open-access
journals. As a graduate student, I began submitting my preprints of my publications
to the arXiv\footnote{\url{https://arxiv.org}}, and later archiving my
(otherwise non-accessible) conference papers on
Figshare\footnote{\url{https://figshare.com}}.

Since then, I have adopted the following policy for my group's research:
\begin{itemize}
    \item All written research products are made openly accessible, either through
    green or gold open-access avenues. Since my field generally lacks recognized,
    fully open-access journals, this objective is typically met by submitting
    preprints to services such as arXiv, engrXiv, or PeerJ Preprints, depending
    on the paper's topic. Conference papers, when not submitted to an open venue,
    are also made openly available through, e.g., Figshare. Where possible, all
    preprints will be released under the Creative Commons Attribution (CC BY)
    license.\footnote{\url{https://creativecommons.org/licenses/by/4.0/legalcode}}
    (I choose, in general, to not follow the hybrid gold open-access model by
    paying a non-fully open journal.)

    \item All new research software developed by my group will be done so openly
    (e.g., on GitHub) and released publicly under a permissive license such as
    the BSD 3-clause license. The Git version-control system will be used to
    track the history of software projects, and releases associated with
    publications or data will be archived (with DOIs) using
    Zenodo\footnote{\url{https://zenodo.org}}.
    In addition, we will make all efforts to describe as many implementation
    details as necessary to reproduce our work.

    \item All data generated through research, when serving as the basis for
    a publication, will be archived publicly and cited appropriately in our
    work. Figures and plotting scripts will be shared via the CC BY license
    and cited in our papers.
\end{itemize}
This policy will be implemented in part by incorporating its content into
funding proposals, for example in Data Management Plans.

In addition to these standards I established for my and my group's research, I
have also gotten involved with and helped initiate some community efforts to
encourage and reward open science practices.

First, after participating in the working group on software credit at the Third
Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE3)\autocite{Katz:2016er}, I joined the FORCE11 Software Citation Working
Group as a cochair, and helped coauthor the resulting Software Citation
Principles\autocite{Smith:2016kt}. One purpose of standardizing software
citation is to help ensure authors\slash developers receive academic credit for
their work in releasing open research software. I also serve on the editorial
boards of the \textit{Journal of Open Research
Software} and \textit{Journal of Open Source Software}\footnote{\url{http://joss.theoj.org/};
\url{http://openresearchsoftware.metajnl.com/}},
which offer publication venues focused solely on open-source research software.

@@

The following list of recommendations are mostly targeted at changing criteria
for promotion and tenure, and performance reviews, to encourage faculty to
practice more open science:
\begin{itemize}
    \item Consider accessibility\slash openness of research products along
    with quantity and ``quality'' in promotion and tenure review.

>>>>

\begin{abstract}
\noindent
This white paper was written as a contribution to the ``Imagining Tomorrow's University:
Rethinking scholarship, education, and institutions for an open, networked era''
workshop, a joint NIH\slash NSF-funded event held 8--9 March 2017 in Rosemont, IL.
In this paper, I present an overview of what I consider open science,
its importance, and how it plays a role in my research agenda.
I also discuss challenges faced in pursuing research openness, and
recommend changes to university leaders to address these barriers.
\end{abstract}

@@

Transforming research communities from traditional, closed environments to open
ones is important for a number of reasons, including (but not necessarily limited
to) the following six. McKiernan et al.\autocite{McKiernan:2016iz} discuss these
and additional benefits for researchers working openly.
Tennant et al.\autocite{Tennant:2016bi} review in detail the benefits
of open-access publications to academics and society.

\paragraph{Accessibility:}
Openness in research ensures that research products,
particularly written output, remain accessible to all. This includes the
research community, funders, policy makers, and the general public. Accessibility
of research products is particularly important for publicly funded
research---since the public paid for the research, the public should have access
to, and be able to benefit from, it.
(This does not prevent innovators or other parties from developing commercial
intellectual property based on the findings, but ensures that the original
discovery, when funded by the public, remains accessible to all.)

\paragraph{Reproducibility:}
Releasing products of research, including software and data, helps enable
reproducibility. This is particularly true for computational science, where a
written description of methods can never describe an approach as completely as
source code. In general, access to research software used to perform a
computational study, or the data from an experimental study, should enable
others to reproduce the findings of the original researchers.
However, open science is a necessary but not sufficient aspect of reproducibility,
as it can be challenging to reproduce or replicate results even with
available research software and data\autocite{Mesnard:2016,Barba:2016ky}.

@@

\paragraph{It's nice:}
In addition to the above benefits, sharing products of research openly is kind
to colleagues and the greater research community, as it prevents people from
wasting time by unnecessarily repeating work.
Like many graduate students, I began working on my dissertation research by
attempting to reimplement another group's method and reproduce some of their
results in a paper. However, this group did not share any source code openly.
As a result, I wasted significant time guessing about minor implementation details
or inputs not discussed in their papers. This could have been avoided by sharing
the source code, which would have allowed me to more quickly move on to new work.
Graduate students and other researchers constantly face similar challenges that
could be avoided by greater openness in research.

@@

Many published journal articles go unread, even in their topical domains.
One study of citation rates found that 27\% of papers published in the natural
sciences and engineering go uncited\autocite{Lariviere:2009}\footnote{Of course,
papers that are read may not be cited, and papers that are cited may not actually
be read.}
Those who do read most papers likely come from research institutions similar to
those of the authors, even if the findings could be impactful beyond these
confines, for example by leading to policy changes or technological solutions
for humanitarian purposes. In part this is due to the challenging technical content,
jargon, and niche topics---but it is also due to a lack of access to the journals
where most research findings reside. (Making the content of these papers actually
understandable or digestible by most is another challenge.)

Considering the high and ever-increasing cost of scholarly journal subscriptions,
research results should not be limited to those with the means to purchase access.
By self-archiving (i.e., green open access) or publishing articles in open-access
journals, researchers can ensure access for all members of society, including
policymakers, funders, members of the media, entrepreneurs, and the general
public---or scientists and engineers in the Global South.

@@

Over the past year and a half (September 2015--February 2017), as I initiated
and built my research group, I worked to crystallize my vision and plan for
performing research openly. The guiding principles I formed were heavily
inspired by Lorena Barba's Reproducibility PI Manifesto\autocite{Barba:2012pi},
the Peer Reviewers' Openness Initiative\autocite{Morey150547}, and others.
However, in addition to these exemplars, my goals are also motivated by the desire
to work more openly than the norm in my field.

My research field includes computational modeling of combustion and fluid dynamics,
chemical kinetics, and associated numerical methods, and lacks reputable open-access
journals. As a graduate student, I began submitting preprints of my publications
to the arXiv\footnote{\url{https://arxiv.org}}, and later archived my
(otherwise non-accessible) conference papers on
Figshare\footnote{\url{https://figshare.com}}.

Since then, I have adopted the following policy for my group's research:
\begin{itemize}
    \item All written research products are made openly accessible, either through
    green or gold open-access avenues. Since my field generally lacks recognized,
    fully open-access journals, this objective is typically met by submitting
    preprints to services such as arXiv, engrXiv, PeerJ Preprints, or Figshare,
    depending on the topic. Conference papers, when not submitted to an open venue,
    are also made openly available. Where possible, all
    preprints will be released under the Creative Commons Attribution (CC BY)
    license.\footnote{\url{https://creativecommons.org/licenses/by/4.0/legalcode}}
    (I choose, in general, to not follow the hybrid gold open-access model by
    paying a non-fully open journal to make a paper accessible.)

    \item All new research software developed by my group will be done so openly
    (e.g., on GitHub) and released publicly under a permissive license, such as
    the BSD 3-clause license. The Git version-control system is used to
    track the history of software projects, and releases associated with
    publications or data are archived (with DOIs) using
    Zenodo\footnote{\url{https://zenodo.org}}.
    In addition, we attempt to describe as many implementation
    details as necessary to reproduce our work.

    \item All data generated through research, when serving as the basis for
    a publication, is archived publicly and cited appropriately in our
    work. Figures and plotting scripts are shared via the CC BY license
    and cited in our papers.
\end{itemize}
This policy will be implemented in part by incorporating its content into
funding proposals, for example in Data Management Plans.

In addition to these standards I established for my group's research, I
have also gotten involved with and helped initiate some community efforts to
encourage and reward open science practices.

First, after participating in the working group on software credit at the Third
Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE3)\autocite{Katz:2016er}, I joined the FORCE11 Software Citation Working
Group as a cochair, and helped coauthor the resulting Software Citation
Principles\autocite{Smith:2016kt}. One purpose of standardizing software
citation is to help ensure authors\slash developers receive academic credit for
their work in releasing open research software. I also serve on the editorial
boards of the \textit{Journal of Open Research
Software}\footnote{\url{http://openresearchsoftware.metajnl.com/}}
and \textit{Journal of Open Source Software}\footnote{\url{http://joss.theoj.org/}},
which offer publication venues focused solely on open-source research software.

@@

The following list of recommendations are mostly targeted at changing criteria
for promotion and tenure, and performance reviews, to encourage faculty to
practice more open science:
\begin{itemize}
    \item Consider accessibility\slash openness of research products along
    with quantity and ``quality'' in promotion and tenure review. Mandate
    self-archiving of publications (i.e., green open access).
